Towards a Machine Learning approach for Aggregate Computing

### 2021-11-04 paper-2021-woa-ac-ml.tex
more on background: move computational model there and enrich the description

In the article, we outline issues, opportunities and methodologies of these integrations, presenting a case study
 where Aggregate Computing is successfully combined with Hysteretic Q-Learning.
%

@@

In this direction, Aggregate Computing~\cite{DBLP:journals/computer/BealPV15} is an innovative solution by which 
 developers can express collective self-organising behaviours through a functional program specification.
%
Indeed, an aggregate program consists of the manipulation of \textit{computational fields}, a distributed
 data structure where each point in the spacetime is associated with the result of a computation
 performed by a node.
%
These operations are defined in \textit{Field-calculus}~\cite{DBLP:conf/coordination/AudritoBDV18} --- the root of Aggregate Computing. 
 This calculus describes a set of minimal constructs to express any spatiotemporal computation.
%
Thanks to the ``composable'' nature of Aggregate Computing, we find common abstractions 
 in collective behaviour specification, and then we encapsulate them in so-called ``building blocks''.
%
With those, different properties have been proven (such as \textit{eventual consistency}~\cite{DBLP:journals/taas/BealVPD17}) and various
 applications have been built (such as swarm robotics, crowds engineering~\cite{DBLP:journals/eaai/CasadeiVAPD21}).

@@

\outline{Probably, for a "short" paper, this section can be avoided.}
\outline{walkthrogh traditional engineering way to design CAS}
\outline{walkthrogh Aggregate Computing}
\subsection{Field calcus}
\todos{I think it can be helpful to introduce fields calculus through ScaFi...}
\subsection{Machine Learning}
\suggestions{I do not know, perhaps it is better to avoid this }
\outline{walkthrogh Machine learning approach applied in MARL}
\outline{current limitation (application-specific, not large scale, centralised in some cases,..)}

@@

\todo[inline, color=green]{"application" pesprective: building block, framework, aggregate computing to guide learning}
\todo[inline, color=green]{learning perspective: homogeneous team learning}
\todo[inline, color=green]{learning modality perspective: offline through simulation, offline with local reward, offline then online, full-online}
\todo[inline, color=green]{method perspective: Supervised, reinforcement, multi-agent reinforcement}

@@

\section{Aggregate Computing with Reinforcement}
In this section, we describe our first result in Aggregate Computing and Machine learning combination.
% 
Hence we briefly describe how nodes compute to follow the global aggregate specification,
 how Reinforcement Learning can be injected in the computation loop,
 and what kind of Reinforcement Learning can be used with Aggregate Computing.

\subsection{Computatational Model}
An aggregate program execution consists of a continuously node-point execution of a \round{}, 
 that is the atomical part of aggregate computing.
%
Each node can interact only with-in its neighbourhoods sharing the \export{}, 
 namely the result produced by the evaluation of a \round{}.

>>>>

In the article, we outline issues, opportunities and methodologies of this integration, presenting a case study
 where Aggregate Computing is successfully combined with Hysteretic Q-Learning.
%

@@

In this direction, Aggregate Computing~\cite{DBLP:journals/computer/BealPV15} is an innovative solution by which 
 developers can express collective self-organising behaviours through a functional program specification.
%
Indeed, an aggregate program (i.e. a program written with Aggregate Computing) consists of the manipulation of \textit{computational fields}, a distributed
 data structure where each point in the spacetime is associated with the result of a computation
 performed by a node.
%
These operations are defined in \textit{field-calculus}~\cite{DBLP:conf/coordination/AudritoBDV18} --- the root of Aggregate Computing. 
 This calculus describes a set of minimal constructs to express any spatiotemporal computation.
%
Thanks to the ``composable'' nature of Aggregate Computing, we find common abstractions 
 in collective behaviour specification, and then we encapsulate them in so-called ``building blocks''.
%
With those, different properties have been proven (such as \textit{eventual consistency}~\cite{DBLP:journals/taas/BealVPD17}) and various
 applications have been built (such as swarm robotics, crowds engineering~\cite{DBLP:journals/eaai/CasadeiVAPD21}).

@@

Engineering this kind of system is a thriving area, various approaches have been proposed.
 First attempts leverage so-called \textit{local-to-global} techniques, where the collective behaviour emerges from local behaviour description, 
 usually inspired by nature~\cite{DBLP:journals/swarm/BrambillaFBD13}. 
%
Though, a natural way to describe this kind of problem is by using an \textit{collective} view of the system.
 And it is the idea used by the \textit{global-to-local} techniques~\cite{DBLP:journals/jlap/ViroliBDACP19,DBLP:journals/scp/AlrahmanNL20, DBLP:conf/cbse/BuresGHKKP13}, where the behaviours are designed \textit{top-down},
 and then the local behaviour is generated automatically. 
% 
Also ``meet-in-the-middle''~\cite{DBLP:journals/computer/PinciroliB16} approaches exist, where
 a part of the system is still expressed bottom-up (device-centric) whereas a part is expressed top-down (swarm-centric)

In our works, we are focused on Aggregate Computing~\cite{DBLP:journals/computer/BealPV15}, global-to-local approach where
 the self-organising collective behaviour is expressed in terms of manipulation of \textit{computational fields}.

The latter could be seen as the extension of electromagnetic fields to computer science since it basically consists in a 
 map from any device to its computational value (i.e. the result of a computation).

The computational field manipulation is ruled by \textit{field calculus}, which define the constructs needed 
 any for express spatio-temporal computation~\cite{DBLP:conf/coordination/AudritoBDV18}.

The novelty of Aggregate Computing consists in its composable (functional-like) program definition: using basic constructs
 we identify the main patterns (\textit{building-blocks}) that could be reusable at the application level.

Crafting opportunistically these building-blocks is possible to verify prominent properties such as eventual consistency~\cite{DBLP:journals/taas/BealVPD17} (i.e. the ensamble behaviour is independent from the underlying network details)
 and self-stabilisation~\cite{DBLP:journals/tomacs/ViroliABDP18} (i.e. the ability of a system to recover from arbitrary changes).

Aggregate Computing is studied in various simulated scenario (thanks to Alchemist, a meta-simulator for pervasive systems), rangin from smart cities to crodws, thaks to programming language and toolchain that support 
 this paradigm, such as ScaFi, FCCP and Protelis.

In the next sections, we show the field calculus operators that are essential to understand the case study presented and then
 we underline how each node computes in order to follow the aggregate program speficiation 
\subsection{Field Calculus operators}

\subsection{Computational Model}
An aggregate program execution consists of a continuously node-point execution of a \round{}, 
 that is the atomical part of aggregate computing.
%
Each node can interact only with-in its neighbourhoods sharing the \export{}, 
 namely the result produced by the evaluation of a \round{}.

\outline{Probably, for a "short" paper, this section can be avoided.}
\outline{walkthrogh traditional engineering way to design CAS}
\outline{walkthrogh Aggregate Computing}
\subsection{Machine Learning}
\suggestions{I do not know, perhaps it is better to avoid this }
\outline{walkthrogh Machine learning approach applied in MARL}
\outline{current limitation (application-specific, not large scale, centralised in some cases,..)}

@@

@@

\section{Aggregate Computing with Reinforcement}
In this section, we describe our first result in Aggregate Computing and Machine learning combination.
% 
Hence we briefly describe how Reinforcement Learning can be injected in the computation loop,
 and what kind of Reinforcement Learning can be used with Aggregate Computing.
